A spiral galaxy's mass distribution uncovered through lensing and dynamics

### 2015-09-14 docs/results_JAMwithNFW_v1.01.tex
Wrote section on future work and removed some TO DOs.

\paragraph{Modelling.} The full set of fit parameters is $(\Upsilon_{I,*},R_\text{S,NFW},v_{200},\beta)$ [TO DO: $\beta$ or $\beta_0$???] [TO DO: Einheitlich $\Upsilon_{I,*}$????], where $\beta$ is the constant velocity anisotropy parameter (see Section \ref{sec:model_JAM}). We will investigate this parameter space with a MCMC \citep{emcee} and use priors for the halo parameters to guide the fit to a realistic NFW halo shape.
\\\citet{Dutton10} give a relation for halo vs. stellar mass for late-type galaxies. For an Chabrier IMF stellar population they found
\begin{equation*}
y = y_0 \left( \frac{x}{x_0} \right)^\alpha \left[ \frac 12 + \frac 12 \left(\frac{x}{x_0} \right)^\gamma \right]^{(\beta - \alpha)/\gamma},
\end{equation*}
where $x= m_*$ [$M_\odot h^{-2}$] is the stellar mass, $y = \frac{\langle M_{200}\rangle}{m_*}$ and $\langle M_{200}\rangle$ [$M_\odot h^{-2}$] is the mean halo mass. The parameters for the mean and $2\sigma$ error curves for this relation are$\alpha = -0.5\pm0.15$, $\beta = 0$, $\gamma = 1.0$, $\log_{10} x_0 = 10.4$, $\log_{10} {y_0}_{-0.24}^{+0.28}$ \citet{Dutton10}. Using the stellar mass estimate for J1331 from \citet{SWELLSI} $m_* = (1.06 \pm 0.25) \cdot 10^{11} M_\odot$ for the Chabrier IMF estimate [TO DO: introduce somewhere], we find ${v_{200}} = (202_{-33}^{+44})_{-13}^{+12}$. The first of the two quoted errors is due to the $2\sigma$ scatter in the relation by \citet{Dutton10}. The second error is the propagated error due to the uncertainty in the stellar mass. We use this as a rough estimate for the halo of J1331 and as Gaussian prior on $v_{200}$, 
\begin{equation*}
p(v_{200}) = \mathscr{N}(200 \text{km/s} ,40 \text{km/s}).
\end{equation*}
\\We also use the concentration vs. halo mass relation by \citet{Maccio08} in Equation \ref{eq:Maccio08} as a prior on the concentration, i.e.
\begin{equation*}
p(\log c_{200} \mid v_{200}) = \mathscr{N}\left(\langle \log c_{200} \rangle (M_{200}) \mid 0.105 \right).
\end{equation*}
\\For the velocity anisotropy parameter $\beta$ we will again employ a uniform prior 
\begin{equation*}
p(\beta) = \mathscr{U}(-0.5,+0.5)
\end{equation*}
to exclude very unrealistic anisotropies. The full prior used is then
\begin{eqnarray*}
&&p(\Upsilon_{I,*},R_\text{S,NFW},v_{200},\beta) \\
&&= \frac{1}{\ln\left( 10 R_{s,NFW}\right)} p(\log c_{200} \mid v_{200}) \cdot p(v_{200}) \cdot p(\beta), 
\end{eqnarray*}
where the factor $1/\ln\left( 10 R_{s,NFW}\right)$ is the Jacobian of the transformation from the halo parameters $(v_{200},\log c_{200})$ [TO DO: introduce concentration] to $(v_{200},R_{s,NFW})$.
\\We restrict the $v_\text{rms}$ fit to a region $R' \lesssim 3.5$ arcsec, approximately within the effective half-light radius $R_\text{eff} = 2.6$ arcsec  [TO DO: make sure R' everywhere ???]. We also include the Einstein mass $M(<R_\text{ein}$ with a 10\% error as an additional constraint in the fit. [TO DO: Didn't include the results from the Moster 2010 relation, because they where quite close to the Dutton 2010 relation anyway. Or should I?]

\paragraph{Result.} Figure \ref{fig:modelB4_triangle} shows the posterior probability distribution of the fit sampled with an MCMC. Overplotted are also the priors used to constrain the NFW halo. The mean parameters are summarized in Table \ref{tab:modelB4_bestfit}. We find that the best fit NFW halo strives to be more massive and with a higher concentration (due to a smaller $R_\text{s,NFW}$) than  proposed by the prior. At the same time the model prefers again very negative velocity anisotropies: The sample with the highest probability is pegged at the lower limit of the unifrom prior. Both, the concentrated halo and the low anisotropy are needed to reproduce the central dip of the $v_\text{rms}$ curve. There are slight covariances between $\Upsilon_{I,*}$,  $R_\text{s,NFW}$ and $\beta_z$: The smaller the dark matter contribution in the center (i.e. the larger $\Upsilon_{I,*}$ [TO DO: make sure, mass-to-light ratio is everywhere named properly]), the less concentrated is the halo (i.e. the larger $R_\text{s,NFW}$), the more velocity anisotropy is needed to reproduce the central dip. As the effect of $v_{200}$ is mostly at larger radii, this parameter doesn't show any covariances - but is also mostly constrained by the prior. The $v_\text{rms}$ curve corresponding to the mean best fit values from Table \ref{tab:modelB4_bestfit} is shown in the first panel of Figure \ref{fig:modelB4_vrot}. Figure \ref{fig:modelB4_models} illustrates the range of parameters fitting the data according to the full posterior probability function. The model fits the $v_\text{rms}$ data in the inner regions quite well (Figure \ref{fig:modelB4_vrms}) and is also consistent with the Einstein mass (see Figure \ref{fig:modelB4_enclMass}). The extrapolation of the model to larger radii however overestimates the data, does not exhibit a drop around $\sim 6$ arcsec at all and seems to be therefore overall too massive.\\

@@

\begin{figure*}
\centering
\includegraphics[width=0.9\linewidth]{fig/B4_contour_plot_short.ps}
\caption{Posterior probability distribution sampled with MCMC (red dots and histograms) for a JAM model with NFW halo, parametrized by $R_\text{s,NFW}$ and $v_\text{200}$, a stellar mass distribution generated from the I-band MGE in Table \ref{tab:MGEF814W} and a constant stellar mass-to-light ratio $\Upsilon_{I,*}$, and constant velocity anisotropy $\beta_z$. Shown are also the priors used for J1331's NFW halo, $\mathscr{N}(200 \text{km/s},40 \text{km/s})$ (green) and the concentration vs. halo mass relation by \citet{Maccio08} from Equation \ref{eq:Maccio08} in terms of $v_{200}$ vs. $R_\text{s,NFW}$ with $1\sigma$ scatter (blue). The MCMC samples are color coded according to their probability (darker red for higher probability); the sample point with the highest probability is marked by a black diamond. The black cross is the mean of the distribution and the ellipses are derived from the covariance of matrix of the sample set and correspond approximately to $1\sigma$ (black solid ellipse) and $2\sigma$ (black dotted ellipse). The histograms of the marginalized 1D distributions are overplotted by the mean (black solid lines) and $1\sigma$ error (black dashed lines), whose values are also quoted in the figure and in Table \ref{tab:modelB4_bestfit}. The grey diamonds mark a random sub-selection of 12 samples; the corresponding models are shown in Figure \ref{fig:modelB4_models}. [TO DO: Rename $\beta_0$ to $\beta_z$, rename $R_\text{s,NFW}$ to $r_\text{s,NFW}$, rename $\Upsilon_*$ to $\Upsilon_{I,*}$]}
\label{fig:modelB4_triangle}
\end{figure*}

\begin{table*}
\centering
\begin{tabular}{llrcl}
\hline
stellar I-band mass-to-light ratio & $\Upsilon_{I,*}$ & 4.2 & $\pm$ & 0.2\\
velocity anisotropy & $\beta_z$ & -0.4 & $\pm$ & 0.1 \\
NFW halo scale length & $R_\text{s,NFW}$ [kpc] & 40 & $\pm$ & 20\\
NFW halo virial velocity & $v_{200}$ [km/s] & 240 & $\pm$ & 40\\
NFW halo concentration & $c_{200}$ & 8 & $\pm$ & 2 \\
NFW halo mass & $M_{200}$ [$10^{12} M_\odot$] & 5 & $\pm$ & 2\\
\hline
\end{tabular}
\caption{Summary of the best fit parameters of the JAM model with NFW halo from the MCMC exploration in Figure \ref{fig:modelB4_triangle}. The halo mass and concentration are calculated from the the best fit $R_\text{s,NFW}$ and $v_{200}$.}
\label{tab:modelB4_bestfit}
\end{table*}

\begin{figure*}
\centering
\begin{subfigure}{\textwidth}
  \centering
  \includegraphics[width=0.8\linewidth]{fig/B4_rms_error_curves.ps}
  \caption{Comparison of the $v_\text{rms}$ data with the best fit JAM models.}
  \label{fig:modelB4_vrms}
\end{subfigure}
\begin{subfigure}{.48\textwidth}
  \centering
  \includegraphics[width=0.9\linewidth]{fig/B4_jam_profiles_errors_short_vcirc.ps}
  \caption{Circular velocity curve.}
  \label{fig:modelB4_vcirc}
\end{subfigure}%
\hspace{.02\textwidth}
\begin{subfigure}{.48\textwidth}
  \centering
  \includegraphics[width=0.9\linewidth]{fig/B4_jam_profiles_errors_short_projmass.ps}
  \caption{Enclosed mass profile.}
  \label{fig:modelB4_enclMass}
\end{subfigure}
\caption{Best fit JAM model including a NFW halo and velocity anisotropy with parameters given in Table \ref{tab:modelB4_bestfit}. The 12 lines shown correspond to the 12 models randomly drawn from the posterior probability distribution and marked as grey diamonds in Figure \ref{fig:modelB4_triangle}. Overplotted are the Einstein radius (red dotted line) and the effective half-light radius (black dotted line). The blue dashed line marks the radius within which the data and model were fitted. \emph{Panel a)} shows the comparison of the symmetrized $v_\text{rms}$ data (solid blue points) with the best fit JAM models including a NFW halo (red solid lines). Also shown is the non-symmetrized data at larger radii (open blue dots) and an extrapolation of the best fit models, using the same model parameters but the I-band surface brightness MGE for the outer regions of J1331 derived from the Ellipse model (red dashed lines). \emph{Panel b)} shows the circular velocity curve of the total mass (green), and separately the contribution of the stellar mass (yellow, again generated from the MGE in Table \ref{tab:F814WMGE}) and dark matter (grey). [TO DO: Is this really the spherical $r$? I would think it is actually the cylindrical $R$.] \emph{Panel c)} shows the corresponding enclosed mass profile. Overplotted is also the Einstein mass at the Einstein radius with a 10\% error, which was used in the fit as an additional constraint.[TO DO: radius benennungen sind irgendwie verrutscht]}
\label{fig:modelB4_models}
\end{figure*}

>>>>

\paragraph{Modelling.} The full set of fit parameters is $(\Upsilon_\text{I,*},R_\text{S,NFW},v_{200},\beta_z)$ [TO DO: $\beta$ or $\beta_0$???], where $\beta_z$ is the constant velocity anisotropy parameter (see Section \ref{sec:model_JAM}). We will investigate this parameter space with a MCMC\footnote{The Python code package for \emph{emcee}, a Monte Carlo Markov Chain implementation by \citet{emcee} is available online at \url{http://dan.iel.fm/emcee/current/}. The version from October 2013 was used in this work.} \citep{emcee} and use priors for the halo parameters to guide the fit to a realistic NFW halo shape.
\\\citet{Dutton10} give a relation for halo vs. stellar mass for late-type galaxies. For an Chabrier IMF stellar population they found
\begin{equation*}
y = y_0 \left( \frac{x}{x_0} \right)^\alpha \left[ \frac 12 + \frac 12 \left(\frac{x}{x_0} \right)^\gamma \right]^{(\beta - \alpha)/\gamma},
\end{equation*}
where $x= m_*$ [$M_\odot h^{-2}$] is the stellar mass, $y = \frac{\langle M_{200}\rangle}{m_*}$ and $\langle M_{200}\rangle$ [$M_\odot h^{-2}$] is the mean halo mass. The parameters for the mean and $2\sigma$ error curves for this relation are$\alpha = -0.5\pm0.15$, $\beta = 0$, $\gamma = 1.0$, $\log_{10} x_0 = 10.4$, $\log_{10} {y_0}_{-0.24}^{+0.28}$ \citet{Dutton10}. Using the stellar mass estimate for J1331 from \citet{SWELLSI} $m_* = (1.06 \pm 0.25) \cdot 10^{11} M_\odot$ for the Chabrier IMF estimate [TO DO: introduce somewhere], we find ${v_{200}} = (202_{-33}^{+44})_{-13}^{+12}$. The first of the two quoted errors is due to the $2\sigma$ scatter in the relation by \citet{Dutton10}. The second error is the propagated error due to the uncertainty in the stellar mass. We use this as a rough estimate for the halo of J1331 and as Gaussian prior on $v_{200}$, 
\begin{equation*}
p(v_{200}) = \mathscr{N}(200 \text{km/s} ,40 \text{km/s}).
\end{equation*}
\\We also use the concentration vs. halo mass relation by \citet{Maccio08} in Equation \ref{eq:Maccio08} as a prior on the concentration, i.e.
\begin{equation*}
p(\log c_{200} \mid v_{200}) = \mathscr{N}\left(\langle \log c_{200} \rangle (M_{200}) \mid 0.105 \right).
\end{equation*}
\\For the velocity anisotropy parameter $\beta$ we will again employ a uniform prior 
\begin{equation*}
p(\beta) = \mathscr{U}(-0.5,+0.5)
\end{equation*}
to exclude very unrealistic anisotropies. The full prior used is then
\begin{eqnarray*}
&&p(\Upsilon_\text{I,*},R_\text{S,NFW},v_{200},\beta) \\
&&= \frac{1}{\ln\left( 10 R_{s,NFW}\right)} p(\log c_{200} \mid v_{200}) \cdot p(v_{200}) \cdot p(\beta), 
\end{eqnarray*}
where the factor $1/\ln\left( 10 R_{s,NFW}\right)$ is the Jacobian of the transformation from the halo parameters $(v_{200},\log c_{200})$ [TO DO: introduce concentration] to $(v_{200},R_{s,NFW})$.
\\We restrict the $v_\text{rms}$ fit to a region $R' \lesssim 3.5$ arcsec, approximately within the effective half-light radius $R_\text{eff} = 2.6$ arcsec  [TO DO: make sure R' everywhere ???]. We also include the Einstein mass $M(<R_\text{ein}$ with a 10\% error as an additional constraint in the fit. [TO DO: Didn't include the results from the Moster 2010 relation, because they where quite close to the Dutton 2010 relation anyway. Or should I?]

\paragraph{Result.} Figure \ref{fig:modelB4_triangle} shows the posterior probability distribution of the fit sampled with an MCMC. Overplotted are also the priors used to constrain the NFW halo. The mean parameters are summarized in Table \ref{tab:modelB4_bestfit}. We find that the best fit NFW halo strives to be more massive and with a higher concentration (due to a smaller $R_\text{s,NFW}$) than  proposed by the prior. At the same time the model prefers again very negative velocity anisotropies: The sample with the highest probability is pegged at the lower limit of the unifrom prior. Both, the concentrated halo and the low anisotropy are needed to reproduce the central dip of the $v_\text{rms}$ curve. There are slight covariances between $\Upsilon_\text{I,*}$,  $R_\text{s,NFW}$ and $\beta_z$: The smaller the dark matter contribution in the center (i.e. the larger $\Upsilon_\text{I,*}$), the less concentrated is the halo (i.e. the larger $R_\text{s,NFW}$), the more velocity anisotropy is needed to reproduce the central dip. As the effect of $v_{200}$ is mostly at larger radii, this parameter doesn't show any covariances - but is also mostly constrained by the prior. The $v_\text{rms}$ curve corresponding to the mean best fit values from Table \ref{tab:modelB4_bestfit} is shown in the first panel of Figure \ref{fig:modelB4_vrot}. Figure \ref{fig:modelB4_models} illustrates the range of parameters fitting the data according to the full posterior probability function. The model fits the $v_\text{rms}$ data in the inner regions quite well (Figure \ref{fig:modelB4_vrms}) and is also consistent with the Einstein mass (see Figure \ref{fig:modelB4_enclMass}). The extrapolation of the model to larger radii however overestimates the data, does not exhibit a drop around $\sim 6$ arcsec at all and seems to be therefore overall too massive.\\

@@

\begin{figure*}
\centering
\includegraphics[width=0.9\linewidth]{fig/B4_contour_plot_short.ps}
\caption{Posterior probability distribution sampled with MCMC (red dots and histograms) for a JAM model with NFW halo, parametrized by $R_\text{s,NFW}$ and $v_\text{200}$, a stellar mass distribution generated from the I-band MGE in Table \ref{tab:MGEF814W} and a constant stellar mass-to-light ratio $\Upsilon_\text{I,*}$, and constant velocity anisotropy $\beta_z$. Shown are also the priors used for J1331's NFW halo, $\mathscr{N}(200 \text{km/s},40 \text{km/s})$ (green) and the concentration vs. halo mass relation by \citet{Maccio08} from Equation \ref{eq:Maccio08} in terms of $v_{200}$ vs. $R_\text{s,NFW}$ with $1\sigma$ scatter (blue). The MCMC samples are color coded according to their probability (darker red for higher probability); the sample point with the highest probability is marked by a black diamond. The black cross is the mean of the distribution and the ellipses are derived from the covariance of matrix of the sample set and correspond approximately to $1\sigma$ (black solid ellipse) and $2\sigma$ (black dotted ellipse). The histograms of the marginalized 1D distributions are overplotted by the mean (black solid lines) and $1\sigma$ error (black dashed lines), whose values are also quoted in the figure and in Table \ref{tab:modelB4_bestfit}. The grey diamonds mark a random sub-selection of 12 samples; the corresponding models are shown in Figure \ref{fig:modelB4_models}. [TO DO: Rename $\beta_0$ to $\beta_z$, rename $R_\text{s,NFW}$ to $r_\text{s,NFW}$, rename $\Upsilon_*$ to $\Upsilon_\text{I,*}$]}
\label{fig:modelB4_triangle}
\end{figure*}

\begin{table*}
\centering
\begin{tabular}{llrcl}
\hline
stellar I-band mass-to-light ratio & $\Upsilon_\text{I,*}$ & 4.2 & $\pm$ & 0.2\\
velocity anisotropy & $\beta_z$ & -0.4 & $\pm$ & 0.1 \\
NFW halo scale length & $R_\text{s,NFW}$ [kpc] & 40 & $\pm$ & 20\\
NFW halo virial velocity & $v_{200}$ [km/s] & 240 & $\pm$ & 40\\
NFW halo concentration & $c_{200}$ & 8 & $\pm$ & 2 \\
NFW halo mass & $M_{200}$ [$10^{12} M_\odot$] & 5 & $\pm$ & 2\\
\hline
\end{tabular}
\caption{Summary of the best fit parameters of the JAM model with NFW halo from the MCMC exploration in Figure \ref{fig:modelB4_triangle}. The halo mass and concentration are calculated from the the best fit $R_\text{s,NFW}$ and $v_{200}$.}
\label{tab:modelB4_bestfit}
\end{table*}

\begin{figure*}
\centering
\begin{subfigure}{\textwidth}
  \centering
  \includegraphics[width=0.8\linewidth]{fig/B4_rms_error_curves.ps}
  \caption{Comparison of the $v_\text{rms}$ data with the best fit JAM models.}
  \label{fig:modelB4_vrms}
\end{subfigure}
\begin{subfigure}{.48\textwidth}
  \centering
  \includegraphics[width=0.9\linewidth]{fig/B4_jam_profiles_errors_short_vcirc.ps}
  \caption{Circular velocity curve.}
  \label{fig:modelB4_vcirc}
\end{subfigure}%
\hspace{.02\textwidth}
\begin{subfigure}{.48\textwidth}
  \centering
  \includegraphics[width=0.9\linewidth]{fig/B4_jam_profiles_errors_short_projmass.ps}
  \caption{Enclosed mass profile.}
  \label{fig:modelB4_enclMass}
\end{subfigure}
\caption{Best fit JAM model including a NFW halo and velocity anisotropy with parameters given in Table \ref{tab:modelB4_bestfit}. The 12 lines shown correspond to the 12 models randomly drawn from the posterior probability distribution and marked as grey diamonds in Figure \ref{fig:modelB4_triangle}. Overplotted are the Einstein radius (red dotted line) and the effective half-light radius (black dotted line). The blue dashed line marks the radius within which the data and model were fitted. \emph{Panel a)} shows the comparison of the symmetrized $v_\text{rms}$ data (solid blue points) with the best fit JAM models including a NFW halo (red solid lines). Also shown is the non-symmetrized data at larger radii (open blue dots) and an extrapolation of the best fit models, using the same model parameters but the I-band surface brightness MGE for the outer regions of J1331 derived from the Ellipse model (red dashed lines). \emph{Panel b)} shows the circular velocity curve of the total mass (green), and separately the contribution of the stellar mass (yellow, again generated from the MGE in Table \ref{tab:MGEF814W}) and dark matter (grey). [TO DO: Is this really the spherical $r$? I would think it is actually the cylindrical $R$.] \emph{Panel c)} shows the corresponding enclosed mass profile. Overplotted is also the Einstein mass at the Einstein radius with a 10\% error, which was used in the fit as an additional constraint.[TO DO: radius benennungen sind irgendwie verrutscht]}
\label{fig:modelB4_models}
\end{figure*}
